---
title: ""
output:
  html_document:
    df_print: paged
---

```{r}
library(ggplot2)
library(dplyr)
library(plotly)
```
Importing dataset
```{r}
df <- read.csv('BCC.csv')
head(df)

```
Exploring data structure
```{r}
str(df)
```

Deleted id
```{r}
df <- select(df, -id)
```

Calculate Principal Component and delete outcome 'diagnosis'

```{r}
pr.out = prcomp(select(df,-diagnosis) , scale = TRUE)
pr.out
```

Explore structure
```{r}
str(pr.out)
```

Summary Principle components
```{r}
summary(pr.out)

```
```{r}
#not include standard deviations 
pr.out$rotation
```
```{r}
#As showed the equation, sum of squared loadings = 1
#Should be 1 
apply (pr.out$rotation, 2, function(x)
  sum(x ^ 2))

```
```{r}
#Should use ggplot
plot(pr.out)
```
```{r}
#Better to use ggplot
# %>% t translate to matrix
# %>%  as.data.frame() translate to dataframe
pc_dat <- summary(pr.out)$importance %>% t %>%  as.data.frame()
class(pc_dat)

pc_dat
```
Creating new row
```{r}
pc_dat$PCs <- rownames(pc_dat)
pc_dat
```
```{r}
#convert to chr to factor
pc_dat$PCs <- factor(pc_dat$PC, levels = c("PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10", "PC11","PC12","PC13","PC14","PC15","PC16","PC17","PC18","PC19","PC20"))
pc_dat
```

Graph relationship between Variance Explained vs Principle of Component
```{r warning=FALSE}
# `Standard deviation`^2 = variance
# + geom_point = highest points
pl <- ggplot(data = pc_dat, aes(x=PCs, y=`Standard deviation`^2))+
  geom_bar(stat="identity", aes(fill=PCs))+
  geom_point()+
  geom_line(aes(group = 1), color="blue")+
  ylab("Variance Explained")+
  xlab("Principal Components")+
  theme_bw()
ggplotly(pl)
```

Graph relationship between Proportion of Variance Explained vs Principle of Component

```{r warning=FALSE}
pl <- ggplot(data = pc_dat, aes(x=PCs, y=`Proportion of Variance`))+
  geom_bar(stat="identity", aes(fill=PCs))+
  geom_point()+
  geom_line(aes(group = 1), color="blue")+
  ylab("Proportion of Variance Explained")+
  xlab("Principal Components")+
  theme_bw()
ggplotly(pl)

```
```{r}
#PC1 explain 85% of variance of data
#Let see loading for each variable 
#sort min -> max
pr.out$rotation[,1] %>% sort

```

Cumulative Proportion of variance accounted by each component 

```{r warning=FALSE}
pl <- ggplot(data = pc_dat, aes(x=PCs, y=`Cumulative Proportion`))+
  geom_bar(stat="identity", aes(fill=PCs))+
  geom_point()+
  geom_smooth(aes(group = 1), formula = y~poly(x,3), color="blue", method="lm", se = F)+
  ylab("Cumulative Proportion of Variance Explained")+
  xlab("Principal Components")+
  theme_bw()
ggplotly(pl)
```
```{r}
#The scale=0 argument to biplot() ensures that the arrows are scaled to represent the loadings; other values for scale give slightly different biplots with different interpretations
biplot (pr.out , scale = 0, cex = .6) 
```
```{r}
library(pca3d)
```

2D graph with PC1 AND PC2 no show.ellipses
```{r}
pca2d(pr.out, group = df$diagnosis, biplot = T, show.labels = F, legend="topleft", radius=2 )
```

2D graph with PC1 AND PC2  show.ellipses
```{r}
#show grouping ellipses
pca2d(pr.out, group = df$diagnosis, biplot = T, biplot.vars=3, show.labels = F, legend="topleft", show.ellipses = T, show.group.labels = T, radius=2)
```

3d graph with PC1 PC2 AND PC3
```{r}
pca3d(pr.out, group = df$diagnosis, biplot = T, biplot.vars=3, show.labels = F, legend="topleft", show.ellipses = T, show.group.labels = T, radius=2)
```
Hierarchical clustering
```{r}
df.h <- select(df,-diagnosis)
rownames(df.h) <- paste(1:nrow(df.h), df$diagnosis, sep="_") #for checking with cluster result
```
```{r}
hc <- hclust(dist(df.h, method = "manhattan"), method = "ward.D2")
#dist() method = "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski"
#hclust() method = "ward.D", "ward.D2", "single", "complete", "average" (= UPGMA), "mcquitty" (= WPGMA), "median" (= WPGMC) or "centroid" (= UPGMC).
```
```{r}
plot(hc) 
```
```{r}
library(dendextend)
```
```{r}
dend <- df.h %>%  scale %>%  dist(method = "euclidean") %>% hclust(method = "complete") %>% as.dendrogram
dend %>% plot
```

```{r}
dend %>%set("labels_cex", 1) %>% set("labels_col", value = c("red", "blue"), k=2) %>% set("branches_k_color", k = 2) %>%  plot %>% abline(h = 12, lty = 2)
```
```{r}
#we can keep series of adjustments in variable
dend <- dend %>% set("labels_cex", 0.5) %>% set("labels_col", value = c("red", "blue"), k=2) %>% set("branches_k_color", k = 2) 

dend %>%  plot 
dend %>%  abline(h = 12, lty = 2)
dend %>%  rect.dendrogram(k=2, border = 8, lty = 5, lwd = 2)
```
```{r}
dend2 <- df.h %>%  scale %>%  dist(method = "euclidean") %>% hclust(method = "average") %>% as.dendrogram
dend2 %>% plot

```
```{r}
dend2 %>%set("labels_cex", 1) %>% set("labels_col", value = c("red", "blue"), k=2) %>% set("branches_k_color", k = 2) %>%  plot %>% abline(h = 12, lty = 2)
```
```{r echo=FALSE}
# library(ggplot2)
# 
# ggd <- as.ggdend(dend)
# 
# ggplot(ggd)+theme_bw() 
# ```
# ```{r}
# ggplot(ggd, horiz = T)+theme_bw()
# ```
# ```{r}
# #circle
# ggplot(ggd, labels = F)+scale_y_reverse(expand = c(0.2, 0))+coord_polar(theta="x")
```

K-means clustering


We will demonstrate with Breast cancer data 
```{r}
#If using all 4 predictor variables
cl <- kmeans(df.h, centers=3) #centers equivalent to k
cl

```
```{r}
table(cl$cluster, df$diagnosis)


```
<p style="color: red">Answer: I have more variables than sample points. The clustering itself has no problems with the p>n situation, however the visualization internally uses princomp (which is incapable of handling p>n) to plot the similarity space projection.</p>
```{r}
#Error more variables than sample points and the principal component analysis that is being done is failing.
# library(cluster)
# clusplot(df.h, cl$cluster, color=T, shade=T, labels=0,lines=0)

```




